Analytic Doppler Imaging

### 2019-08-16 tex/ms.tex
twospec working

%
\begin{center}
    \begin{longtable}{cll}
    \caption{Notation used in this paper} 
    \label{tab:notation} \\
    %
    \toprule
    \multicolumn{1}{c}{\textbf{Symbol}} &
    \multicolumn{1}{c}{\textbf{Description}} &
    \multicolumn{1}{c}{\textbf{Type}} \\
    \toprule
    \endhead
    %
    \endfoot
    %
    \endlastfoot
    %
    $a$                             & lowercase italic & scalar\\
    $A$                             & uppercase italic & scalar\\
    $\alpha$                        & lowercase italic Greek & scalar\\
    $\bvec{a}$                      & lowercase boldface & column vector\\
    $\bvec{A}$                      & uppercase boldface & matrix\\
    $\boldsymbol{\alpha}^\top$      & lowercase boldface Greek & basis row vector
\end{longtable}
\end{center}

%
\section{The problem}
\label{sec:the_problem}
%
Let $I(\lnlam, \x, t, \Dargs)$ be the specific 
intensity observed 
at log wavelength $\lnlam \equiv \ln\lambda$ and at sky-projected 
Cartesian position $\x = (x, y, z)$ on the surface of the 
star at time $t$, where
$\Dargs$ is a set of arbitrary parameters of the Doppler field.
We may express this intensity as
%
\begin{align}
    \label{eq:intro:Ixi}
    I(\lnlam, \x, t, \Dargs) &= 
        I\Big(\lnlam_0 + \D(\x, \Dargs), \x, t\Big)
\end{align}
%
where $\lnlam_0$ is the log wavelength in the rest frame and $\D$ is
the Doppler shift in log wavelength space:
%
\begin{align}
    \label{eq:intro:D}
    \D(\x, \Dargs) 
        &=
        \frac{1}{2}\ln\left( 
            \frac{1 + \beta(\x, \Dargs)}{1 - \beta(\x, 
            \Dargs)} 
        \right)
\end{align}
%
where $\beta = v(\x, \Dargs) / c$ is the ratio of the 
radial velocity at a point on the surface of the star to the speed of light.
In keeping with the literature, we take positive values of $v$ (and
$\D$) to mean redshifts.

A common approach to computing the Doppler-shifted spectrum is to
evaluate the spectrum at the rest frame wavelength $\lnlam_0$
and interpolate back to the grid in $\lnlam$. This is practical when
computing the spectrum at a single \emph{point} on the surface, but not
ideal when one is interested in the \emph{integral} over the visible
surface of the star $\Surf$, which is typically all we can observe:
%
\begin{align}
    \label{eq:intro:F}
    F(\lnlam, t, \Dargs) 
        &\equiv
        \iint\limits_{\Surf(\x)}
                I(\lnlam, \x, t, \Dargs)
        \mathrm{d}{\Surf(\x)}
        \quad ,
\end{align}
%
The difficulty in solving Equation~(\ref{eq:intro:F}) stems from the fact
that $I(\lnlam, \x, t, \Dargs)$ is difficult to write down in 
closed form, given
the non-linearity of the Doppler shift.
The standard approach to solving this integral is therefore
to discretize the surface of the star with a fine grid, evaluate the
Doppler-shifted spectrum in each cell, and sum over the spatial axes
to approximate the integral. Depending on the resolution of the grid,
this is either numerically inaccurate or computationally inefficient 
(and often both).

@@

The observed spectrum is a complex function
of spatial, spectral, temporal, and velocity terms. The goal in this
section is to deconvolve each of these terms to make solving the integral
in Equation~(\ref{eq:intro:F}) easier.
%
The first thing we will do is to express Equation~(\ref{eq:intro:Ixi})
as a convolution:
%
\begin{align}
    \label{eq:deconv:convolution}
    I(\lnlam, \x, t, \Dargs) &= 
        I(\lnlam_0, \x, t) 
        * 
        \delta\big(\lnlam_0 + \D(\x, \Dargs)\big)
\end{align}
%
where $\delta$ is the
Dirac delta function and
$*$ denotes the linear convolution operator, defined for
two arbitrary functions $g$ and $h$ as the integral
%
\begin{align}
    \label{eq:deconv:convolution_def}
    (g * h)(t) \equiv \int_{-\infty}^\infty f(\tau) h(t - \tau) d\tau
\end{align}
%
for some independent coordinate $t$ and dummy parameter $\tau$.
%
The convolution of $I(\lnlam_0)$ with a delta function
has the effect of shifting the spectrum by an amount $\D$, returning
a function of the shifted (observed) log wavelength, 
$\lnlam = \lnlam_0 + \D$.

Next, we expand the spatial dependence of the
specific intensity at the rest frame wavelength
in terms of spherical harmonics on the unit disk:
%
\begin{align}
    \label{eq:deconv:Ixi0}
    I(\lnlam_0, \x, t) 
        &=
        \sum_{l=0}^\infty\sum_{m=-l}^{l} a_{lm}(\lnlam_0, t) Y_{lm}(\x)
    \quad ,
\end{align}
%
where $Y_{lm}(\x)$ is a spherical harmonic on the projected disk
and $a_{lm}(\lnlam_0, t)$ is the corresponding spherical harmonic 
coefficient at log wavelength in the rest frame $\lnlam_0$ and time $t$. For 
convenience, we may write this equation in vector form:
%
\begin{align}
    \label{eq:deconv:Ivec}
    I(\lnlam_0, \x, t) &=
    \ylmbasis(\x) \,
    \alm(\lnlam_0, t)
    \quad ,
\end{align}
%
where
%
\begin{align}
    \label{eq:deconv:alm}
    \alm(\lnlam_0, t) \equiv
\Big( 
    a_{0,0}(\lnlam_0, t) \quad\quad\quad\quad\quad\quad 
    a_{1,-1}(\lnlam_0, t) \quad\quad\quad\quad\quad\quad 
    a_{1,0}(\lnlam_0, t) \quad\quad\quad\quad\quad\quad
    a_{1,1}(\lnlam_0, t) \quad\quad\quad\quad\quad\quad 
    ... 
\Big)^\top
\end{align}
%
is the vector of spherical harmonic coefficients and
%
\begin{align}
    \ylmbasis(\x) \equiv 
\Big( 
    Y_{0,0}(\x) \quad\quad\quad\quad\quad\quad 
    Y_{1,-1}(\x) \quad\quad\quad\quad\quad\quad 
    Y_{1,0}(\x) \quad\quad\quad\quad\quad\quad 
    Y_{1,1}(\x) \quad\quad\quad\quad\quad\quad 
    ... 
\Big)
\end{align}
%
is the corresponding vector of spherical harmonics. We may further
decompose our expression by linearizing the time dependence of the
spherical harmonic coefficients:
%
\begin{align}
    \label{eq:deconv:R}
    \alm(\lnlam_0, t) = \R(t) \, \azero(\lnlam_0)
    \quad ,
\end{align}
%
where $\azero(\lnlam_0) = \bvec{a}(\lnlam_0, t=t_0)$ 
for some reference time $t_0$.
For rigid body rotation, the result is exact and $\R(t)$ is the Wigner 
rotation matrix for real spherical harmonics 
\citep[e.g.][]{AlvarezCollado1989}, which is implicitly a 
function of the inclination, obliquity, and rotation period of the star.
%In the case that other processes such as differential rotation or spot 
%evolution are significant
%over the course of an observation, Equation~(\ref{eq:deconv:R}) can be
%made to hold approximately for some effective rotation matrix 
%$\R(t)$; we discuss this later.

@@

\subsection{Computing the kernel $\kT$}
\label{sec:kT}
%
In the case that the star's rotation is rigid (i.e., differential rotation
is negligible) and other effects such as convective blueshift
may be ignored, the radial velocity at any point on the surface is 
simply proportional to the distance to the axis of rotation. Without loss
of generality, if we assume the axis of rotation lies along the $y-z$ plane,
we have
%
\begin{align}
    \beta(\x, \Dargs) = \frac{\omega \sin i \, x}{c}
\end{align}
%
and
%
\begin{align}
    \label{eq:kT:D}
    \D(\x, \Dargs) &= 
        \frac{1}{2}\ln\left( 
            \dfrac{1 + \dfrac{\omega \sin i \, x}{c}}
                 {1 - \dfrac{\omega \sin i \, x}{c}}
        \right)
    \quad ,
\end{align}
%
where the parameters $\Dargs$ of the Doppler field are the
angular velocity, $\omega$, and the inclination of the star with
respect to $\hat{y}$, $i$.

@@

%
\begin{align}
    \label{eq:norm:f}
    \bvec{f}
    &=
    \frac{
        \Doppler
        \,
        \azero
    } {
        \bvec{b}
    }
    \quad,
\end{align}
%
where $\bvec{b}$ is the baseline and the vector division is
taken elementwise. When $\azero$ is given by Equation~(\ref{eq:inverse:azero}),
the baseline $\bvec{b}$ is simply the total (wavelength-integrated) flux
at each epoch, repeated for every element in each spectrum. 
Based on \citet{Luger2019}, this can be computed as a purely linear operation 
on the spherical harmonic vector $\bvec{y}$:
%
\begin{align}
    \label{eq:norm:b}
    \bvec{b}
    &=
    %
    \begin{pmatrix}
        \bvec{r}^\top \bvec{A_1} \bvec{R}(t_0) \\
        \cdots \\
        \bvec{r}^\top \bvec{A_1} \bvec{R}(t_0) \\
        \bvec{r}^\top \bvec{A_1} \bvec{R}(t_1) \\
        \cdots \\
        \bvec{r}^\top \bvec{A_1} \bvec{R}(t_1) \\
        \cdots \\
        \bvec{r}^\top \bvec{A_1} \bvec{R}(t_{M-1}) \\
        \cdots \\
        \bvec{r}^\top \bvec{A_1} \bvec{R}(t_{M-1})
    \end{pmatrix}
    \bvec{y}
    %
    \nonumber \\[0.5em]
    &\equiv
    \bvec{B} \, \bvec{y}
\end{align}
%
where $\bvec{r}$ is given by Equation~(18) in \citet{Luger2019},
$\bvec{A_1}$ is given in Appendix B of \citet{Luger2019}, and $\bvec{R}(t)$
is the rotation matrix defined previously.

Given this new equation, we must change how we approach the inverse problem
(\S\ref{sec:inverse}). When solving for the spectrum given the spherical 
harmonic vector (\S\ref{sec:solve_s}), we need only make the replacements
$\bvec{f} \rightarrow \nicefrac{\bvec{f}}{\bvec{B}\, \bvec{y}}$
and 
$\boldsymbol{\Sigma}_\mathbf{f} \rightarrow \nicefrac{\boldsymbol{\Sigma}_\mathbf{f}}{\bvec{B}\, \bvec{y}}$ 
in Equations~(\ref{eq:inverse:shat}) and (\ref{eq:inverse:scov}).
However, when solving for the spherical harmonic vector given the spectrum
(\ref{sec:solve_y}), the problem is no longer linear in $\bvec{y}$, as the
vector of interest appears in both the numerator and the denominator:
%
\begin{align}
    \label{eq:norm:fy}
    \bvec{f}
    &=
    \frac{
        \bvec{D}
        \,
        \bvec{S}
        \,
        \bvec{y}
    }{
        \bvec{B}
        \,
        \bvec{y}
    }
    \quad.
\end{align}
%
While we can solve for $\bvec{y}$ using a non-linear optimizer, it is
often the case that the denominator in Equation~(\ref{eq:norm:fy}) is close
to unity; i.e., the total flux from the star typically varies by only a
few percent. When this is the case, it is straightforward to linearize
this equation.
Following the notation from \S\ref{sec:solve_y}, we may therefore write
%
\begin{align}
    \bvec{f}
    &=
    \frac{
        \bvec{D} \, \bvec{s_0}
        \,
        + 
        \,
        \bvec{D}\bvec{S_1}
        \,
        \bvec{y_1}
    }{
        \bvec{B_0}
        \,
        +
        \bvec{B_1}
        \,
        \bvec{y_1}
    }
    \quad,
\end{align}
%
where $\bvec{B_0} = \bvec{1}$ is the first column of $\bvec{B}$
and $\bvec{B}_\bvec{1}$ denotes the
matrix composed from the remaining $N - 1$ columns of $\bvec{B}$.
Expanding the expression above about 
$\bvec{B_1} \bvec{y_1} = \bvec{0}$, we have
%
\begin{align}
    \bvec{f}
    &\approx
    \left(
        \bvec{D} \, \bvec{s_0}
        \,
        + 
        \,
        \bvec{D}\bvec{S_1}
        \,
        \bvec{y_1}
    \right)
    \circ
    \left(
        \bvec{1}
        \,
        -
        \bvec{B_1}
        \,
        \bvec{y_1}
    \right)
    \nonumber \\
    &=
    \left(
        \bvec{D}\bvec{S_1}
        \,
        -
        \bvec{C}
    \right)
    \bvec{y_1}
    \,
    +
    \,
    \bvec{D} \, \bvec{s_0}
    \quad,
\end{align}
%
where $\circ$ denotes an elementwise product and $\bvec{C}$ is a matrix
constructed by multiplying the $i^\mathrm{th}$ row of $\bvec{B_1}$ by the
$i^\mathrm{th}$ element of $\bvec{D}\bvec{s_0}$.
%
This equation is now
linear in $\bvec{y}_\bvec{1}$, so we may solve the least-squares problem
as before:
%
\begin{align}
    \label{eq:norm:yhat}
    \bvec{\hat{y}_1} &= 
    \boldsymbol{\Sigma}_\mathbf{\hat{y}_1}
    \left(
        \left(\bvec{D}\bvec{S_1} - \bvec{C}\right)^\top
        {\boldsymbol{\Sigma}_\mathbf{f}}^{-1}
        (\bvec{f} - \bvec{D}\bvec{s_0})
        +
        {\boldsymbol{\Lambda}_{\mathbf{y}_\bvec{1}}}^{-1} \boldsymbol{\mu}_{\mathbf{y}_\bvec{1}}
    \right)
    \quad,
\end{align}
%
with covariance
%
\begin{align}
    \label{eq:norm:ycov}
    \boldsymbol{\Sigma}_\mathbf{\hat{y}_1} &= 
    \left(
        \left(\bvec{D}\bvec{S_1} - \bvec{C}\right)^\top
        {\boldsymbol{\Sigma}_\mathbf{f}}^{-1}
        \left(\bvec{D}\bvec{S_1} - \bvec{C}\right)
        +
        {\boldsymbol{\Lambda}_{\mathbf{y}_1}}^{-1}
    \right)^{-1}
    \quad.
\end{align}
%
Note, importantly, that these equations are valid only in the limit that the
change in the total flux from the star is small. Nevertheless, even in the 
general case, we find that solving this linear problem yields a good
starting point for a non-linear optimization, which we revisit in more
detail below.

>>>>

%
\begin{center}
    \begin{longtable}{cll}
    \caption{Notation used in this paper} 
    \label{tab:notation} \\
    %
    \toprule
    \multicolumn{1}{c}{\textbf{Symbol}} &
    \multicolumn{1}{c}{\textbf{Description}} &
    \multicolumn{1}{c}{\textbf{Reference}} \\
    \toprule
    \endhead
    %
    \endfoot
    %
    \endlastfoot
    %
    $a_{lm}$    & spectral spherical harmonic coefficient               & Equation~(\ref{eq:deconv:Ixi0})\\
    $\alm$      & vector of spectral spherical harmonic coefficients    & Equation~(\ref{eq:deconv:alm})\\
    $\bvec{a_0}$& $\alm$ at reference time $t = t_0$                    & Equation~(\ref{eq:deconv:R})\\
    $c$         & speed of light                                        & Equation~(\ref{eq:kT:beta})\\
    $\Curve$    & curve of constant radial velocity                     & Equation~(\ref{eq:deconv:kT})\\
    $\Dargs$    & parameters of the Doppler field                       & \S\ref{sec:the_problem}\\
    $\D$        & logarithmic Doppler shift                             & Equation~(\ref{eq:the_problem:D})\\
    $F$         & flux                                                  & Equation~(\ref{eq:the_problem:F})\\
    $i$         & inclination                                           & Equation~(\ref{eq:kT:beta})\\
    $I$         & specific intensity                                    & Equation~(\ref{eq:the_problem:Ixi})\\
    $l$         & spherical harmonic degree                             & \\
    $m$         & spherical harmonic order                              & \\
    $\bvec{R}$  & Wigner rotation matrix                                & Equation~(\ref{eq:deconv:R})\\
    $\Surf$     & projected visible area of the star                    & Equation~(\ref{eq:the_problem:F})\\
    $t$         & time                                                  & \S\ref{sec:the_problem}\\
    $\x$        & Cartesian position on the sky                         & \S\ref{sec:the_problem}\\
    $y_{lm}$    & spherical harmonic coefficient                        & \S\ref{sec:inverse}\\
    $Y_{lm}$    & spherical harmonic                                    & Equation~(\ref{eq:deconv:Ixi0})\\
    %
    %
    $\beta$     & relativistic parameter                                & Equation~(\ref{eq:the_problem:D})\\
    $\delta$    & Dirac delta function                                  & Equation~(\ref{eq:deconv:convolution})\\
    $\kT$       & convolution kernel basis vector                       & Equation~(\ref{eq:deconv:kT})\\
    $\lambda$   & wavelength (in nm)                                    & \S\ref{sec:the_problem}\\
    $\lnlam$    & $\ln\lambda$ (observer frame)                         & \S\ref{sec:the_problem}\\
    $\lnlam_0$  & $\ln\lambda$ (rest frame)                             & \S\ref{sec:the_problem}\\
    $\ylmbasis$ & spherical harmonic basis vector                       & Equation~(\ref{eq:deconv:ylmbasis})\\
    $\omega$    & angular velocity                                      & Equation~(\ref{eq:kT:beta})\\
    %
    %
    $*$         & convolution                                           & Equation~(\ref{eq:deconv:convolution_def})\\
    $\circ$     & element-wise product                                  & Equation~(\ref{eq:norm:fapprox})
\end{longtable}
\end{center}

%
\section{The problem}
\label{sec:the_problem}
%
Let $I(\lnlam, \x, t, \Dargs)$ be the specific 
intensity observed 
at log wavelength $\lnlam \equiv \ln\lambda$ and at sky-projected 
Cartesian position $\x = (x, y, z)$ on the surface of the 
star at time $t$, where
$\Dargs$ is a set of arbitrary parameters of the Doppler field.
We may express this intensity as
%
\begin{align}
    \label{eq:the_problem:Ixi}
    I(\lnlam, \x, t, \Dargs) &= 
        I\Big(\lnlam_0 + \D(\x, \Dargs), \x, t\Big)
\end{align}
%
where $\lnlam_0$ is the log wavelength in the rest frame and $\D$ is
the Doppler shift in log wavelength space:
%
\begin{align}
    \label{eq:the_problem:D}
    \D(\x, \Dargs) 
        &=
        \frac{1}{2}\ln\left( 
            \frac{1 + \beta(\x, \Dargs)}{1 - \beta(\x, 
            \Dargs)} 
        \right)
\end{align}
%
where $\beta = v(\x, \Dargs) / c$ is the ratio of the 
radial velocity at a point on the surface of the star to the speed of light.
In keeping with the literature, we take positive values of $v$ (and
$\D$) to mean redshifts.

A common approach to computing the Doppler-shifted spectrum is to
evaluate the spectrum at the rest frame wavelength $\lnlam_0$
and interpolate back to the grid in $\lnlam$. This is practical when
computing the spectrum at a single \emph{point} on the surface, but not
ideal when one is interested in the \emph{integral} over the visible
surface of the star $\Surf$, which is typically all we can observe:
%
\begin{align}
    \label{eq:the_problem:F}
    F(\lnlam, t, \Dargs) 
        &\equiv
        \iint\limits_{\Surf(\x)}
                I(\lnlam, \x, t, \Dargs)
        \mathrm{d}{\Surf(\x)}
        \quad ,
\end{align}
%
The difficulty in solving Equation~(\ref{eq:the_problem:F}) stems from the fact
that $I(\lnlam, \x, t, \Dargs)$ is difficult to write down in 
closed form, given
the non-linearity of the Doppler shift.
The standard approach to solving this integral is therefore
to discretize the surface of the star with a fine grid, evaluate the
Doppler-shifted spectrum in each cell, and sum over the spatial axes
to approximate the integral. Depending on the resolution of the grid,
this is either numerically inaccurate or computationally inefficient 
(and often both).

@@

The observed spectrum is a complex function
of spatial, spectral, temporal, and velocity terms. The goal in this
section is to deconvolve each of these terms to make solving the integral
in Equation~(\ref{eq:the_problem:F}) easier.
%
The first thing we will do is to express Equation~(\ref{eq:the_problem:Ixi})
as a convolution:
%
\begin{align}
    \label{eq:deconv:convolution}
    I(\lnlam, \x, t, \Dargs) &= 
        I(\lnlam_0, \x, t) 
        * 
        \delta\big(\lnlam_0 + \D(\x, \Dargs)\big)
\end{align}
%
where $\delta$ is the
Dirac delta function and
$*$ denotes the linear convolution operator, defined for
two arbitrary functions $g$ and $h$ as the integral
%
\begin{align}
    \label{eq:deconv:convolution_def}
    (g * h)(t) \equiv \int_{-\infty}^\infty g(\tau) h(t - \tau) d\tau
\end{align}
%
for some independent coordinate $t$ and dummy parameter $\tau$.
%
The convolution of $I(\lnlam_0)$ with a delta function
has the effect of shifting the spectrum by an amount $\D$, returning
a function of the shifted (observed) log wavelength, 
$\lnlam = \lnlam_0 + \D$.

Next, we expand the spatial dependence of the
specific intensity at the rest frame wavelength
in terms of spherical harmonics on the unit disk:
%
\begin{align}
    \label{eq:deconv:Ixi0}
    I(\lnlam_0, \x, t) 
        &=
        \sum_{l=0}^\infty\sum_{m=-l}^{l} a_{lm}(\lnlam_0, t) Y_{lm}(\x)
    \quad ,
\end{align}
%
where $Y_{lm}(\x)$ is a spherical harmonic on the projected disk
and $a_{lm}(\lnlam_0, t)$ is the corresponding spherical harmonic 
coefficient at log wavelength in the rest frame $\lnlam_0$ and time $t$. For 
convenience, we may write this equation in vector form:
%
\begin{align}
    \label{eq:deconv:Ivec}
    I(\lnlam_0, \x, t) &=
    \ylmbasis(\x) \,
    \alm(\lnlam_0, t)
    \quad ,
\end{align}
%
where
%
\begin{align}
    \label{eq:deconv:alm}
    \alm(\lnlam_0, t) \equiv
\Big( 
    a_{0,0}(\lnlam_0, t) \quad\quad\quad\quad\quad\quad 
    a_{1,-1}(\lnlam_0, t) \quad\quad\quad\quad\quad\quad 
    a_{1,0}(\lnlam_0, t) \quad\quad\quad\quad\quad\quad
    a_{1,1}(\lnlam_0, t) \quad\quad\quad\quad\quad\quad 
    ... 
\Big)^\top
\end{align}
%
is the vector of spherical harmonic coefficients and
%
\begin{align}
    \label{eq:deconv:ylmbasis}
    \ylmbasis(\x) \equiv 
\Big( 
    Y_{0,0}(\x) \quad\quad\quad\quad\quad\quad 
    Y_{1,-1}(\x) \quad\quad\quad\quad\quad\quad 
    Y_{1,0}(\x) \quad\quad\quad\quad\quad\quad 
    Y_{1,1}(\x) \quad\quad\quad\quad\quad\quad 
    ... 
\Big)
\end{align}
%
is the corresponding vector of spherical harmonics. We may further
decompose our expression by linearizing the time dependence of the
spherical harmonic coefficients:
%
\begin{align}
    \label{eq:deconv:R}
    \alm(\lnlam_0, t) = \R(t) \, \azero(\lnlam_0)
    \quad ,
\end{align}
%
where $\azero(\lnlam_0) = \bvec{a}(\lnlam_0, t=t_0)$ 
for some reference time $t_0$.
For rigid body rotation, the result is exact and $\R(t)$ is the Wigner 
rotation matrix for real spherical harmonics 
\citep[e.g.][]{AlvarezCollado1989}, which is implicitly a 
function of the inclination, obliquity, and rotation period of the star.
%In the case that other processes such as differential rotation or spot 
%evolution are significant
%over the course of an observation, Equation~(\ref{eq:deconv:R}) can be
%made to hold approximately for some effective rotation matrix 
%$\R(t)$; we discuss this later.

@@

\subsection{Computing the kernel $\kT$}
\label{sec:kT}
%
In the case that the star's rotation is rigid (i.e., differential rotation
is negligible) and other effects such as convective blueshift
may be ignored, the radial velocity at any point on the surface is 
simply proportional to the distance to the axis of rotation. Without loss
of generality, if we assume the axis of rotation lies along the $y-z$ plane,
we have
%
\begin{align}
    \label{eq:kT:beta}
    \beta(\x, \Dargs) = \frac{\omega \sin i \, x}{c}
\end{align}
%
and
%
\begin{align}
    \label{eq:kT:D}
    \D(\x, \Dargs) &= 
        \frac{1}{2}\ln\left( 
            \dfrac{1 + \dfrac{\omega \sin i \, x}{c}}
                 {1 - \dfrac{\omega \sin i \, x}{c}}
        \right)
    \quad ,
\end{align}
%
where the parameters $\Dargs$ of the Doppler field are the
angular velocity, $\omega$, and the inclination of the star with
respect to $\hat{y}$, $i$.

@@

%
\begin{align}
    \label{eq:norm:f}
    \bvec{f}
    &=
    \frac{
        \Doppler
        \,
        \azero
    } {
        \bvec{b}
    }
    \quad,
\end{align}
%
where $\bvec{b}$ is the baseline and the vector division is
performed elementwise. When $\azero$ is given by Equation~(\ref{eq:inverse:azero}),
the baseline $\bvec{b}$ is simply the total (wavelength-integrated) flux
at each epoch, repeated for every element in each spectrum. 
Based on \citet{Luger2019}, this can be computed as a purely linear operation 
on the spherical harmonic vector $\bvec{y}$:
%
\begin{align}
    \label{eq:norm:b}
    \bvec{b}
    &=
    %
    \begin{pmatrix}
        \bvec{r}^\top \bvec{A_1} \bvec{R}(t_0) \\
        \cdots \\
        \bvec{r}^\top \bvec{A_1} \bvec{R}(t_0) \\
        \bvec{r}^\top \bvec{A_1} \bvec{R}(t_1) \\
        \cdots \\
        \bvec{r}^\top \bvec{A_1} \bvec{R}(t_1) \\
        \cdots \\
        \bvec{r}^\top \bvec{A_1} \bvec{R}(t_{M-1}) \\
        \cdots \\
        \bvec{r}^\top \bvec{A_1} \bvec{R}(t_{M-1})
    \end{pmatrix}
    \bvec{y}
    %
    \nonumber \\[0.5em]
    &\equiv
    \bvec{B} \, \bvec{y}
\end{align}
%
where $\bvec{r}$ is given by Equation~(18) in \citet{Luger2019},
$\bvec{A_1}$ is given in Appendix B of \citet{Luger2019}, and $\bvec{R}(t)$
is the rotation matrix defined previously.

Given this new equation, we must change how we approach the inverse problem
(\S\ref{sec:inverse}). When solving for the spectrum given the spherical 
harmonic vector (\S\ref{sec:solve_s}), we need only make the replacements
$\bvec{f} \rightarrow \nicefrac{\bvec{f}}{\bvec{B}\, \bvec{y}}$
and 
$\boldsymbol{\Sigma}_\mathbf{f} \rightarrow \nicefrac{\boldsymbol{\Sigma}_\mathbf{f}}{(\bvec{B}\, \bvec{y})^2}$ 
in Equations~(\ref{eq:inverse:shat}) and (\ref{eq:inverse:scov}).
However, when solving for the spherical harmonic vector given the spectrum
(\ref{sec:solve_y}), the problem is no longer linear in $\bvec{y}$, as the
vector of interest appears in both the numerator and the denominator:
%
\begin{align}
    \label{eq:norm:fy}
    \bvec{f}
    &=
    \frac{
        \bvec{D}
        \,
        \bvec{S}
        \,
        \bvec{y}
    }{
        \bvec{B}
        \,
        \bvec{y}
    }
    \quad.
\end{align}
%
While we can solve for $\bvec{y}$ using a non-linear optimizer, it is
often the case that the denominator in Equation~(\ref{eq:norm:fy}) is close
to unity; i.e., the total flux from the star typically varies by only a
few percent. When this is the case, it is straightforward to linearize
this equation.
Following the notation from \S\ref{sec:solve_y}, we may therefore write
%
\begin{align}
    \bvec{f}
    &=
    \frac{
        \bvec{D} \, \bvec{s_0}
        \,
        + 
        \,
        \bvec{D}\bvec{S_1}
        \,
        \bvec{y_1}
    }{
        \bvec{B_0}
        \,
        +
        \bvec{B_1}
        \,
        \bvec{y_1}
    }
    \quad,
\end{align}
%
where $\bvec{B_0} = \bvec{1}$ is the first column of $\bvec{B}$
and $\bvec{B}_\bvec{1}$ denotes the
matrix composed from the remaining $N - 1$ columns of $\bvec{B}$.
Expanding the expression above about 
$\bvec{B_1} \bvec{y_1} = \bvec{0}$, we have
%
\begin{align}
    \label{eq:norm:fapprox}
    \bvec{f}
    &\approx
    \left(
        \bvec{D} \, \bvec{s_0}
        \,
        + 
        \,
        \bvec{D}\bvec{S_1}
        \,
        \bvec{y_1}
    \right)
    \circ
    \left(
        \bvec{1}
        \,
        -
        \bvec{B_1}
        \,
        \bvec{y_1}
    \right)
    \nonumber \\
    &=
    \left(
        \bvec{D}\bvec{S_1}
        \,
        -
        \bvec{C}
    \right)
    \bvec{y_1}
    \,
    +
    \,
    \bvec{D} \, \bvec{s_0}
    \quad,
\end{align}
%
where $\circ$ denotes an elementwise product and $\bvec{C}$ is a matrix
constructed by multiplying the $i^\mathrm{th}$ row of $\bvec{B_1}$ by the
$i^\mathrm{th}$ element of $\bvec{D}\bvec{s_0}$.
%
This equation is now
linear in $\bvec{y}_\bvec{1}$, so we may solve the least-squares problem
as before:
%
\begin{align}
    \label{eq:norm:yhat}
    \bvec{\hat{y}_1} &= 
    \boldsymbol{\Sigma}_\mathbf{\hat{y}_1}
    \left(
        \left(\bvec{D}\bvec{S_1} - \bvec{C}\right)^\top
        {\boldsymbol{\Sigma}_\mathbf{f}}^{-1}
        (\bvec{f} - \bvec{D}\bvec{s_0})
        +
        {\boldsymbol{\Lambda}_{\mathbf{y}_\bvec{1}}}^{-1} \boldsymbol{\mu}_{\mathbf{y}_\bvec{1}}
    \right)
    \quad,
\end{align}
%
with covariance
%
\begin{align}
    \label{eq:norm:ycov}
    \boldsymbol{\Sigma}_\mathbf{\hat{y}_1} &= 
    \left(
        \left(\bvec{D}\bvec{S_1} - \bvec{C}\right)^\top
        {\boldsymbol{\Sigma}_\mathbf{f}}^{-1}
        \left(\bvec{D}\bvec{S_1} - \bvec{C}\right)
        +
        {\boldsymbol{\Lambda}_{\mathbf{y}_1}}^{-1}
    \right)^{-1}
    \quad.
\end{align}
%
Note, importantly, that these equations are valid only in the limit that the
change in the total flux from the star is small. Nevertheless, even in the 
general case, we find that solving this linear problem yields a good
starting point for a non-linear optimization, which we revisit in more
detail below.
